dynamics

### 2023-02-28 scripts/workshop/dynamics.Rmd
knit it

    dV <- sigma*Ia + sigma*Ib - epsilon*V
    res<-c(dSa,dIa,dRa,dSb,dIb,dRb,dV)
    list(res)
  })}
maxTime <- 300.0 # time is in years - run model for this time
times<-seq(0,maxTime,by=1) # how long we run the model for
# notes on params
# mu =     <- density-independent mortality rate
# beta =       <- contact transmission rate by infected A
# betab =       <- contact transmission rate by infected B
# phi =      <- environmental transmission rate
# lambda =       <- density-dependent growth rate
# alpha =     <- recovery rate
# sigma =    <- shedding rate
# epsilon =       <- viral degradation rate
params<-c(mu_a=1/50,
          mu_b=1/50,
          beta=0.001,
          betab=0.0001,
          phi=0.001,
          lambda=5/3,
          alpha=1/10,
          sigma=1/2,
          epsilon=1/2)  # model parameters

@@

output_n <- as.data.frame(lsoda(xstart_n,times,rm0,params_n)) # tells computer to solve (integrate) equations

@@

```{r}
output_n %>% ggplot(.,aes(x=time))+
  geom_line(aes(y=Sa,col="Susceptible A"))+
  geom_line(aes(y=Ia,col="Infected A"))+
  geom_line(aes(y=Ra,col="Recovered A"))+
  geom_line(aes(y=Sb,col="Susceptible B"))+
  geom_line(aes(y=Ib,col="Infected B"))+
  geom_line(aes(y=Rb,col="Recovered B"))+
  geom_line(aes(y=V,col="Virus"))+
  scale_colour_manual(values = c("red","orange","yellow","green","blue","black","violet"))+
  theme(legend.position = c(0.85,0.5))+
  labs(y="N",x="Time",col="Population")+
  ylim(0,95)+
  ggtitle(paste("2-Species Transmission Model for Ranavirus in an Amphibian Wetland - Neutral - ", reference_eigen))+theme(plot.title = element_text(hjust=0, size=10))
```

@@

output_c <- as.data.frame(lsoda(xstart_n,times,rm0,params_c)) # tells computer to solve (integrate) equations

@@

```{r}
output_c %>% ggplot(.,aes(x=time))+
  geom_line(aes(y=Sa,col="Susceptible A"))+
  geom_line(aes(y=Ia,col="Infected A"))+
  geom_line(aes(y=Ra,col="Recovered A"))+
  geom_line(aes(y=Sb,col="Susceptible B"))+
  geom_line(aes(y=Ib,col="Infected B"))+
  geom_line(aes(y=Rb,col="Recovered B"))+
  geom_line(aes(y=V,col="Virus"))+
  scale_colour_manual(values = c("red","orange","yellow","green","blue","black","violet"))+
  theme(legend.position = c(0.85,0.5))+
  labs(y="N",x="Time",col="Population")+
  ylim(0,95)+
  ggtitle(paste("2-Species Transmission Model for Ranavirus in an Amphibian Wetland - Community - ", community_eigen))+theme(plot.title = element_text(hjust=0, size=10))
```

@@

output_a <- as.data.frame(lsoda(xstart_n,times,rm0,params_a)) # tells computer to solve (integrate) equations

@@

```{r}
output_a %>% ggplot(.,aes(x=time))+
  geom_line(aes(y=Sa,col="Susceptible A"))+
  geom_line(aes(y=Ia,col="Infected A"))+
  geom_line(aes(y=Ra,col="Recovered A"))+
  geom_line(aes(y=Sb,col="Susceptible B"))+
  geom_line(aes(y=Ib,col="Infected B"))+
  geom_line(aes(y=Rb,col="Recovered B"))+
  geom_line(aes(y=V,col="Virus"))+
  scale_colour_manual(values = c("red","orange","yellow","green","blue","black","violet"))+
  theme(legend.position = c(0.85,0.5))+
  labs(y="N",x="Time",col="Population")+
  ylim(0,95)+
  ggtitle(paste("2-Species Transmission Model for Ranavirus in an Amphibian Wetland - Abundance - ", abundance_eigen))+theme(plot.title = element_text(hjust=0, size=10))
```

@@

output_h <- as.data.frame(lsoda(xstart_n,times,rm0,params_h)) # tells computer to solve (integrate) equations

@@

```{r}
output_h %>% ggplot(.,aes(x=time))+
  geom_line(aes(y=Sa,col="Susceptible A"))+
  geom_line(aes(y=Ia,col="Infected A"))+
  geom_line(aes(y=Ra,col="Recovered A"))+
  geom_line(aes(y=Sb,col="Susceptible B"))+
  geom_line(aes(y=Ib,col="Infected B"))+
  geom_line(aes(y=Rb,col="Recovered B"))+
  geom_line(aes(y=V,col="Virus"))+
  scale_colour_manual(values = c("red","orange","yellow","green","blue","black","violet"))+
  theme(legend.position = c(0.85,0.5))+
  labs(y="N",x="Time",col="Population")+
  ylim(0,95)+
  ggtitle(paste("2-Species Transmission Model for Ranavirus in an Amphibian Wetland - halflife - ", halflife_eigen))+theme(plot.title = element_text(hjust=0, size=10))
```

@@

output_x <- as.data.frame(lsoda(xstart_n,times,rm0,params_x)) # tells computer to solve (integrate) equations

@@

```{r}
output_x %>% ggplot(.,aes(x=time))+
  geom_line(aes(y=Sa,col="Susceptible A"))+
  geom_line(aes(y=Ia,col="Infected A"))+
  geom_line(aes(y=Ra,col="Recovered A"))+
  geom_line(aes(y=Sb,col="Susceptible B"))+
  geom_line(aes(y=Ib,col="Infected B"))+
  geom_line(aes(y=Rb,col="Recovered B"))+
  geom_line(aes(y=V,col="Virus"))+
  scale_colour_manual(values = c("red","orange","yellow","green","blue","black","violet"))+
  theme(legend.position = c(0.85,0.5))+
  labs(y="N",x="Time",col="Population")+
  ylim(0,95)+
  ggtitle(paste("2-Species Transmission Model for Ranavirus in an Amphibian Wetland - Combined - ", combined_eigen))+theme(plot.title = element_text(hjust=0, size=10))
```

Make figure with infected peaks from everything
```{r}
output_a %<>% mutate(I = Ia+Ib)
output_c %<>% mutate(I = Ia+Ib)
output_h %<>% mutate(I = Ia+Ib)
output_n %<>% mutate(I = Ia+Ib)
output_x %<>% mutate(I = Ia+Ib)
```

@@

>>>>

    dV <- sigma*Ia + sigma*Ib - epsilon*V
    res<-c(dSa,dIa,dRa,dSb,dIb,dRb,dV)
    list(res)
  })}
#no demography
rm1<-function(t,x,params){
  Sa <- x[1]  
  Ia <- x[2]
  Ra <- x[3]
  Sb <- x[4]  
  Ib <- x[5]
  Rb <- x[6]
  V <- x[7]
  with(as.list(params),{
    dSa <- -beta*Sa*Ia - betab*Sa*Ib - phi*Sa*V
    dIa <- beta*Sa*Ia + betab*Sa*Ib + phi*Sa*V - alpha*Ia
    dRa <- alpha*Ia

    dSb <- -betab*Sb*Ib - beta*Sb*Ia - phi*Sb*V
    dIb <- betab*Sb*Ib + beta*Sb*Ia + phi*Sb*V - alpha*Ib
    dRb <- alpha*Ib

    dV <- sigma*Ia + sigma*Ib - epsilon*V
    res<-c(dSa,dIa,dRa,dSb,dIb,dRb,dV)
    list(res)
  })}

maxTime <- 300.0 # time is in years - run model for this time
times<-seq(0,maxTime,by=1) # how long we run the model for
# notes on params
# mu =     <- density-independent mortality rate
# beta =       <- contact transmission rate by infected A
# betab =       <- contact transmission rate by infected B
# phi =      <- environmental transmission rate
# lambda =       <- density-dependent growth rate
# alpha =     <- recovery rate
# sigma =    <- shedding rate
# epsilon =       <- viral degradation rate
params<-c(mu_a=1/50,
          mu_b=1/50,
          beta=0.001,
          betab=0.0001,
          phi=0.001,
          lambda=5/3,
          alpha=1/10,
          sigma=1/2,
          epsilon=1/2)  # model parameters

@@

output_n <- as.data.frame(lsoda(xstart_n,times,rm0,params_n)) # tells computer to solve (integrate) equations
output_n1<- as.data.frame(lsoda(xstart_n,times,rm1,params_n)) # tells computer to solve (integrate) equations

@@

```{r}
output_n %>% ggplot(.,aes(x=time))+
  geom_line(aes(y=Sa,col="Susceptible A"))+
  geom_line(aes(y=Ia,col="Infected A"))+
  geom_line(aes(y=Ra,col="Recovered A"))+
  geom_line(aes(y=Sb,col="Susceptible B"))+
  geom_line(aes(y=Ib,col="Infected B"))+
  geom_line(aes(y=Rb,col="Recovered B"))+
  geom_line(aes(y=V,col="Virus"))+
  scale_colour_manual(values = c("red","orange","yellow","green","blue","black","violet"))+
  theme(legend.position = c(0.85,0.5))+
  labs(y="N",x="Time",col="Population")+
  ylim(0,95)+
  ggtitle(paste("2-Species Transmission Model for Ranavirus in an Amphibian Wetland - Neutral - ", reference_eigen))+theme(plot.title = element_text(hjust=0, size=10))

output_n1 %>% ggplot(.,aes(x=time))+
  geom_line(aes(y=Sa,col="Susceptible A"))+
  geom_line(aes(y=Ia,col="Infected A"))+
  geom_line(aes(y=Ra,col="Recovered A"))+
  geom_line(aes(y=Sb,col="Susceptible B"))+
  geom_line(aes(y=Ib,col="Infected B"))+
  geom_line(aes(y=Rb,col="Recovered B"))+
  geom_line(aes(y=V,col="Virus"))+
  scale_colour_manual(values = c("red","orange","yellow","green","blue","black","violet"))+
  theme(legend.position = c(0.85,0.5))+
  labs(y="N",x="Time",col="Population")+
  ylim(0,95)+
  ggtitle(paste("2-Species Transmission Model for Ranavirus in an Amphibian Wetland w/o demo - Neutral - ", reference_eigen))+theme(plot.title = element_text(hjust=0, size=10))
```

@@

output_c <- as.data.frame(lsoda(xstart_n,times,rm0,params_c)) # tells computer to solve (integrate) equations
output_c1 <- as.data.frame(lsoda(xstart_n,times,rm1,params_c)) # tells computer to solve (integrate) equations

@@

```{r}
output_c %>% ggplot(.,aes(x=time))+
  geom_line(aes(y=Sa,col="Susceptible A"))+
  geom_line(aes(y=Ia,col="Infected A"))+
  geom_line(aes(y=Ra,col="Recovered A"))+
  geom_line(aes(y=Sb,col="Susceptible B"))+
  geom_line(aes(y=Ib,col="Infected B"))+
  geom_line(aes(y=Rb,col="Recovered B"))+
  geom_line(aes(y=V,col="Virus"))+
  scale_colour_manual(values = c("red","orange","yellow","green","blue","black","violet"))+
  theme(legend.position = c(0.85,0.5))+
  labs(y="N",x="Time",col="Population")+
  ylim(0,95)+
  ggtitle(paste("2-Species Transmission Model for Ranavirus in an Amphibian Wetland - Community - ", community_eigen))+theme(plot.title = element_text(hjust=0, size=10))

output_c1 %>% ggplot(.,aes(x=time))+
  geom_line(aes(y=Sa,col="Susceptible A"))+
  geom_line(aes(y=Ia,col="Infected A"))+
  geom_line(aes(y=Ra,col="Recovered A"))+
  geom_line(aes(y=Sb,col="Susceptible B"))+
  geom_line(aes(y=Ib,col="Infected B"))+
  geom_line(aes(y=Rb,col="Recovered B"))+
  geom_line(aes(y=V,col="Virus"))+
  scale_colour_manual(values = c("red","orange","yellow","green","blue","black","violet"))+
  theme(legend.position = c(0.85,0.5))+
  labs(y="N",x="Time",col="Population")+
  ylim(0,95)+
  ggtitle(paste("2-Species Transmission Model for Ranavirus in an Amphibian Wetland w/o demo - Community - ", community_eigen))+theme(plot.title = element_text(hjust=0, size=10))
```

@@

output_a <- as.data.frame(lsoda(xstart_n,times,rm0,params_a)) # tells computer to solve (integrate) equations
output_a1 <- as.data.frame(lsoda(xstart_n,times,rm1,params_a)) # tells computer to solve (integrate) equations

@@

```{r}
output_a %>% ggplot(.,aes(x=time))+
  geom_line(aes(y=Sa,col="Susceptible A"))+
  geom_line(aes(y=Ia,col="Infected A"))+
  geom_line(aes(y=Ra,col="Recovered A"))+
  geom_line(aes(y=Sb,col="Susceptible B"))+
  geom_line(aes(y=Ib,col="Infected B"))+
  geom_line(aes(y=Rb,col="Recovered B"))+
  geom_line(aes(y=V,col="Virus"))+
  scale_colour_manual(values = c("red","orange","yellow","green","blue","black","violet"))+
  theme(legend.position = c(0.85,0.5))+
  labs(y="N",x="Time",col="Population")+
  ylim(0,95)+
  ggtitle(paste("2-Species Transmission Model for Ranavirus in an Amphibian Wetland - Abundance - ", abundance_eigen))+theme(plot.title = element_text(hjust=0, size=10))

output_a1 %>% ggplot(.,aes(x=time))+
  geom_line(aes(y=Sa,col="Susceptible A"))+
  geom_line(aes(y=Ia,col="Infected A"))+
  geom_line(aes(y=Ra,col="Recovered A"))+
  geom_line(aes(y=Sb,col="Susceptible B"))+
  geom_line(aes(y=Ib,col="Infected B"))+
  geom_line(aes(y=Rb,col="Recovered B"))+
  geom_line(aes(y=V,col="Virus"))+
  scale_colour_manual(values = c("red","orange","yellow","green","blue","black","violet"))+
  theme(legend.position = c(0.85,0.5))+
  labs(y="N",x="Time",col="Population")+
  ylim(0,95)+
  ggtitle(paste("2-Species Transmission Model for Ranavirus in an Amphibian Wetland w/o demo - Abundance - ", abundance_eigen))+theme(plot.title = element_text(hjust=0, size=10))
```

@@

output_h <- as.data.frame(lsoda(xstart_n,times,rm0,params_h)) # tells computer to solve (integrate) equations
output_h1 <- as.data.frame(lsoda(xstart_n,times,rm1,params_h)) # tells computer to solve (integrate) equations

@@

```{r}
output_h %>% ggplot(.,aes(x=time))+
  geom_line(aes(y=Sa,col="Susceptible A"))+
  geom_line(aes(y=Ia,col="Infected A"))+
  geom_line(aes(y=Ra,col="Recovered A"))+
  geom_line(aes(y=Sb,col="Susceptible B"))+
  geom_line(aes(y=Ib,col="Infected B"))+
  geom_line(aes(y=Rb,col="Recovered B"))+
  geom_line(aes(y=V,col="Virus"))+
  scale_colour_manual(values = c("red","orange","yellow","green","blue","black","violet"))+
  theme(legend.position = c(0.85,0.5))+
  labs(y="N",x="Time",col="Population")+
  ylim(0,95)+
  ggtitle(paste("2-Species Transmission Model for Ranavirus in an Amphibian Wetland - halflife - ", halflife_eigen))+theme(plot.title = element_text(hjust=0, size=10))

output_h1 %>% ggplot(.,aes(x=time))+
  geom_line(aes(y=Sa,col="Susceptible A"))+
  geom_line(aes(y=Ia,col="Infected A"))+
  geom_line(aes(y=Ra,col="Recovered A"))+
  geom_line(aes(y=Sb,col="Susceptible B"))+
  geom_line(aes(y=Ib,col="Infected B"))+
  geom_line(aes(y=Rb,col="Recovered B"))+
  geom_line(aes(y=V,col="Virus"))+
  scale_colour_manual(values = c("red","orange","yellow","green","blue","black","violet"))+
  theme(legend.position = c(0.85,0.5))+
  labs(y="N",x="Time",col="Population")+
  ylim(0,95)+
  ggtitle(paste("2-Species Transmission Model for Ranavirus in an Amphibian Wetland w/o demo - halflife - ", halflife_eigen))+theme(plot.title = element_text(hjust=0, size=10))
```

@@

output_x <- as.data.frame(lsoda(xstart_n,times,rm0,params_x)) # tells computer to solve (integrate) equations
output_x1 <- as.data.frame(lsoda(xstart_n,times,rm1,params_x)) # tells computer to solve (integrate) equations

@@

```{r}
output_x %>% ggplot(.,aes(x=time))+
  geom_line(aes(y=Sa,col="Susceptible A"))+
  geom_line(aes(y=Ia,col="Infected A"))+
  geom_line(aes(y=Ra,col="Recovered A"))+
  geom_line(aes(y=Sb,col="Susceptible B"))+
  geom_line(aes(y=Ib,col="Infected B"))+
  geom_line(aes(y=Rb,col="Recovered B"))+
  geom_line(aes(y=V,col="Virus"))+
  scale_colour_manual(values = c("red","orange","yellow","green","blue","black","violet"))+
  theme(legend.position = c(0.85,0.5))+
  labs(y="N",x="Time",col="Population")+
  ylim(0,95)+
  ggtitle(paste("2-Species Transmission Model for Ranavirus in an Amphibian Wetland - Combined - ", combined_eigen))+theme(plot.title = element_text(hjust=0, size=10))

output_x1 %>% ggplot(.,aes(x=time))+
  geom_line(aes(y=Sa,col="Susceptible A"))+
  geom_line(aes(y=Ia,col="Infected A"))+
  geom_line(aes(y=Ra,col="Recovered A"))+
  geom_line(aes(y=Sb,col="Susceptible B"))+
  geom_line(aes(y=Ib,col="Infected B"))+
  geom_line(aes(y=Rb,col="Recovered B"))+
  geom_line(aes(y=V,col="Virus"))+
  scale_colour_manual(values = c("red","orange","yellow","green","blue","black","violet"))+
  theme(legend.position = c(0.85,0.5))+
  labs(y="N",x="Time",col="Population")+
  ylim(0,95)+
  ggtitle(paste("2-Species Transmission Model for Ranavirus in an Amphibian Wetland - Combined - ", combined_eigen))+theme(plot.title = element_text(hjust=0, size=10))
```

Make figure with infected peaks from everything
```{r}
output_a %<>% mutate(I = Ia+Ib)
output_c %<>% mutate(I = Ia+Ib)
output_h %<>% mutate(I = Ia+Ib)
output_n %<>% mutate(I = Ia+Ib)
output_x %<>% mutate(I = Ia+Ib)

output_a1 %<>% mutate(I = Ia+Ib)
output_c1 %<>% mutate(I = Ia+Ib)
output_h1 %<>% mutate(I = Ia+Ib)
output_n1 %<>% mutate(I = Ia+Ib)
output_x1 %<>% mutate(I = Ia+Ib)
```

@@

```{r}
ggplot()+
  geom_line(data=output_x1,mapping = aes(y=I,x=time, color=color_x))+
  scale_color_manual(name="",values = "black")+
  new_scale_color()+
  geom_line(data=output_c1, mapping=aes(y=I,x=time, color=color_c))+
  scale_color_manual(name="",values = "red")+
  new_scale_color()+
  geom_line(output_a1, mapping=aes(y=I,x=time, color=color_a))+
  scale_color_manual(name="",values = "orange")+
  new_scale_color()+
  geom_line(output_h1, mapping=aes(y=I,x=time, color=color_h))+
  scale_color_manual(name="",values = "purple")+
  new_scale_color()+
  geom_line(output_n1, mapping=aes(y=I,x=time, color=color_n))+
  scale_color_manual(name="",values = "forestgreen")+
  labs(y="N",x="Time")+
  ylim(0,30)+
  ggtitle(paste("Total Infected Individuals (2 Species) Over Time Across Varying Conditions w/o demo"))+
  theme_bw()
```
